Tampering Detection In Low-Power Smart Cameras

### 2015-05-19 paper.tex
ripassato abstract e sez 1,2. Manca da sistemare related works

\begin{abstract}
A desirable feature for smart cameras is the ability to autonomously detect any tampering event that would prevent a clear view over the monitored scene. No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious events (e.g., the device displacement), these have to be promptly detected to possibly activate countermeasures. Tampering detection becomes particularly challenging in battery-powered cameras, where it is not possible to acquire images at video-like frame-rates, nor use sophisticated image-analysis algorithms.

We here introduce a tampering-detection algorithm that has been specifically designed for low-power smart cameras: the algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme. Any frame yielding anomalous values of the indicators are detected as a tampering events. Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to substantially improve the detection performance for camera displacements. Our experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one per minute, and at a very small computational complexity. %and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene

@@

% What is the problem?
When cameras operate outdoor and in harsh environments, dust, rain drops or snow flakes might lie on the camera lens resulting in blurry pictures (as in Fig. \ref{fig:pioggia}), or in partial occlusions of the scene (as in Fig. \ref{fig:neve}). Similarly, other intentional attacks like displacing the camera, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures (surely for monitoring purposes). We refer to these events/attacks as tampering.

In some cases, tampering is easy to detect, e.g., when the camera integrity is affected and the device goes out-of-order. However, in some other situations, the device might not be affected itself, even though tampering prevents the correct interpretation of the scene or of small important details (e.g., the identification of licence plates).

% Why is it interesting and important?
The early detection of tampering events/attacks is clearly an essential feature for surveillance systems \gi{REFERENCES}, where cameras are expected to autonomously detect any tampering event/attack and promptly report alerts. Surveillance cameras are typically connected to the power supply and acquire and process images at normal frame-rates (e.g. around 30fps) \gi{REFERENCES?}.

\gi{Qui ST pu\'o aggiungere qualche esempio di applicazione qui o menzionare dispositivi di riferimento (con tanto di link a datasheet). Magari aggiungendo che le batterie recenti permettono operativit\'a di due anni a qeusti regimi.}
We here consider low-power cameras, which are battery powered and that have to operate at very low frame-rates (e.g. one frame every minute or more) and that do not have much computational power or memory availability. As an example, consider wireless multimedia sensor networks (WMSN) \cite{akyildiz2007survey} where nodes are smart cameras performing very simple operations, wirelessy connected to other devices such that they can acquire and transmit images at regular time interval or upon requests. In these WMSN, notes are often not connected to the power supply and have to operate battery-powered, possibly implementing energy harvesting mechanisms \gi{REFERENCES, magari riprescando quelle del paper di TIM}. Even though these devices are not employed in critical surveillance applications, low-power smart cameras are becoming popular thanks to their low price which makes them suited for wide monitoring campaign. Tampering detection is therefore very important also in low-power smart cameras, because tampered frames are not informative and as such should not be transmitted over the network, also because operations like radio-module activation are quite energy demanding. Thus, low-power smart cameras have to promptly report tampering attack without raising many false alerts since these lead unnecessary data-transmissions and to energy consumption.

% Why is it hard? (E.g., why do naive approaches fail?)
Tampering detection in low-power smart cameras is much more challenging than in conventional surveillance cameras. Beside computational aspects -- such as the number of operations per pixels allowed and the memory availability -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), and the acquired sequence does not evolve smoothly. This prevents the use of learned background models and the analysis of foreground variations \gi{REFERENCE su BKGR SUBTRACTION}. When dynamic environments are acquired at low frame rates, like the example depicted in Fig. \ref{fig:buenosAires}, two consecutive frames might depict very different scenes because of changes in the depicted content and in the light conditions: smart cameras have to correctly distinguish between these \emph{normal} changes and changes due to camera displacement or blur/defocus.

@@

%What are the key components of my approach and results? Also include any specific limitations.
We here propose an algorithm to detect both camera defocus and of displacement, as stated in Section \ref{sec:probForm}. The algorithm relies on simple indicators that are computed with a low computational complexity (Section \ref{subsec:Indicators}) and leverages an outlier detection technique to detect frames yielding anomalous indicator values as related to a tampering event (Section \ref{subsec:MonitoringScheme}). In particular, we perform a preliminarily partitioning of the scene into nonoverlapping regions, and separately compute and monitore indicators over different regions. Scene partitioning is defined by clustering feature vectors gathering the indicators in different pixels, thus performing a very coarse segmentation of the image (Section \ref{subsec:Segmentation}). Regions are computed during the initial configuration of the algorithm and as such do not implies any computational overhead w.r.t. operating on the whole image. Our experiments in Section \ref{sec:experiments} show that regions substantially improve the detection of camera displacements, while are not beneficial for detecting blur/defocus. Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.

@@

\subsection{Related Works}\label{subsec:relWorks}
% Why hasn't it been solved before? (Or, what's wrong with previous proposed solutions? How does mine differ?)
The literature concerning tampering detection is mostly focused on video surveillance applications and operates at frame-rates around (15- 30 fps) \gi{Adriano, \'e vero?}. Background models are typically leveraged to identify defocus and occlusions; in particular  ~\cite{aksay2007camera} performs detect defocus by analyzing the wavelet domain of each frame and performs histogram comparison for detecting occlusions, while camera displacements are not considered. A background-subtraction technique is employed in~\cite{saglam2009real} to identify defocus, occlusions, and displacements. \gi{FIXME Comparison in the Fourier domain for defocus detection, histogram comparison for occlusion detection, comparison between current background and delayed background for displacement detection}. Background subtraction is also used in \cite{gil2007automatic} to detect defocus, occlusions, and displacements. \gi{FIXME Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection}.
In contrast, no background models are used in \cite{alippi2010detecting}, where a sequential monitoring scheme based on a change-detection test is employed to detect changes in the average gradient energy of each frame to detect defocus due to external disturbances on the camera lens.  
\cite{ribnick2006real}: comparison between frames belonging to a buffer in order to find high values of dissimilarity, associated to tampering.
\cite{kryjak2012fpga}: implementation in a FPGA of a solution based on background modeling, histograms comparisons, edges comparisons.
\cite{harasse2004automated}: tampering detection inside a moving vehicle; uses background subtraction methods in order to identify defocus, occlusions, and displacements. Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection.
\cite{tsesmelis2013tamper}: monitoring of the number of key points extracted by SURF in order to detect defocus events, partition in blocks and HOG descriptors matching for each block in order to detect occlusions. These types of solutions requires a lot of computations

@@

Let us denote by $z_t$ the frame acquired at time $t$, which we model as
\begin{equation}
\label{eq:observationModel}
z_t(x)=\mathcal{D}_t[y_t](x) \quad \forall x \in X
\end{equation}
where $\mathcal{D}_t$ denotes the degradation operator that transforms the original image $y_t$ in the frame $z_t$ which we acquire; $X \subset \mathbb{Z}$ denotes the regular pixels grid and $x$ is a vector indicating the pixel coordinates. As far as there is no tampering attacks/events,
\begin{equation}
\label{eq:no_tampering}
\mathcal{D}_t[y_t](x) = y_t(x) + \eta_t(x) \quad \forall x \in X
\end{equation}
where $\eta_t$ is a random variable accounting for all the noise sources (thermal, quantization, photon-counting). In normal conditions, all the images $y_t$ (thus also the frames $z_t$) might be different in content but are acquired from the same viewpoint and the same camera orientation.

When at time $\tau^*$ an external disturbance over the camera lens introduces a defocus, the image $y_t$ is blurred by spatially variant operator:
\begin{equation}
\label{eq:model_defocus}
\mathcal{D}_t[y_t](x) = \int_{\mathcal{X}}y(s)h_t(x,s)ds\, + \eta_t(x) \quad \forall x \in X, t \leq \tau^*
\end{equation}
%where  $\mathcal{B}_t$ corresponds to a linear operator
%\begin{equation}
%\label{eq:blur}
%\mathcal{B}[y](x) = \int_{\mathcal{X}}y(s)h(x,s)ds\,,
%\end{equation}
where $h_t(x,\cdot) > 0$ is the point-spread function at pixel $x \in X$.

%\begin{equation}
%\label{eq:blur_convolution}
%\mathcal{B}[y](x) = \int_{\mathcal{X}}y(s)h(s-x)ds,
%\end{equation}
%
%\begin{equation}
%\label{eq:blur_multi}
%z_t(x)=\mathcal{D}_t[y_t](x) = \mathcal{B}_t[y_t](x) + \eta(x), \qquad x \in \mathcal{X}.
%\end{equation}
A camera displacement at frame $\tau^*$ is modeled as 
\begin{equation}
\label{eq:model_displacement}
z_t(x)  = \left\{ \begin{array}{rcl}
y_t(x) + \eta(x) & \mbox{per} & t < T^* \\
w_t(x) + \eta(x) & \mbox{per} & t \geqslant T^*
\end{array}\right. ,
\end{equation}
where $w_t$ is acquired from a different viewpoint and/or camera orientation than $y_t$.

The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ and detect frames at time $\tau^*$ that correspond to a tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement}.

>>>>

\begin{abstract}
A desirable feature for smart cameras is the ability to autonomously detect any tampering event/attack that would prevent a clear view over the monitored scene. No matter whether tampering is due to atmospheric phenomena (e.g., few rain drops over the camera lens) or to malicious attack (e.g., the device displacement), these have to be promptly detected to possibly activate countermeasures. Tampering detection becomes particularly challenging in battery-powered cameras, where it is not possible to acquire images at video-like frame-rates, nor use sophisticated image-analysis algorithms.

We here introduce a tampering-detection algorithm that has been specifically designed for low-power smart cameras: the algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme. Any frame yielding anomalous indicator is detected as a tampering attempt. Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to substantially improve the detection of camera displacements. Our experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, and at a very small computational complexity. %and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene

@@

% What is the problem?
When cameras operate outdoor and in harsh environments, dust, rain drops or snow flakes might lie on the camera lens resulting in blurry pictures, as in Fig. \ref{fig:pioggia}, or in partial occlusions of the scene, as in Fig. \ref{fig:neve}. Similarly, other intentional attacks like displacing the camera, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures, that would be surely useless for monitoring purposes. We refer to these events/attacks as tampering. In some cases, tampering is easy to detect, e.g., when the camera integrity is affected and the device goes out-of-order. However, in many other situations, namely when the device is not physically damaged but the correct interpretation of the scene or of small important details is prevented (e.g., the identification of licence plates), tampering detection is not straightforward and requires image-analysis algorithms.

% Why is it interesting and important?
The early detection of tampering events/attacks is clearly essential in surveillance systems \gi{Adriano, add REFERENCES}, where cameras are expected to autonomously detect any tampering event/attack and promptly report alerts. Surveillance cameras are typically connected to the power supply and acquire and process images at normal frame-rates (e.g. around few frames per second). In these conditions, several algorithms have been presented in the literature \gi{Adriano add REFERENCES}.

\gi{Qui ST pu\'o aggiungere qualche esempio di applicazione qui o menzionare dispositivi di riferimento (con tanto di link a datasheet). Magari aggiungendo che le batterie recenti permettono operativit\'a di due anni a questi regimi.}
We here consider low-power cameras, which are battery powered and that have to operate at very low frame-rates (e.g. one frame every minute or more) and which have a constrained computational power or memory availability. As an example, consider wireless multimedia sensor networks (WMSN) \cite{akyildiz2007survey} where nodes are wirelessy connected smart-cameras, that can acquire and transmit images at regular time interval or upon requests. In these WMSN, notes are often not connected to the power supply and have to operate battery-powered, possibly implementing energy harvesting mechanisms \gi{REFERENCES, magari riprescando quelle del paper di TIM}. Even though these devices are not employed in critical surveillance applications, low-power smart cameras are becoming popular in distributed monitoring of wide environments due to their low price \gi{possiamo mettere qualche REFERENCE}. Tampering detection is very important in low-power smart cameras, also considering that tampered frames should not be transmitted over the network, saving energy-demanding operations like the activation of the radio module.

% Why is it hard? (E.g., why do naive approaches fail?)
Tampering detection in low-power smart cameras is much more challenging than in conventional surveillance cameras. Beside computational aspects -- such as the number of operations per pixels allowed and the memory availability -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), and the acquired sequence does not evolve smoothly. This prevents the use of learned background models and the analysis of foreground variations \gi{REFERENCE su BKGR SUBTRACTION}. When dynamic environments are acquired at low frame rates, like the example depicted in Fig. \ref{fig:buenosAires}, two consecutive frames might be very different because of changes in the scene and in the light conditions: smart cameras have to correctly distinguish between these \emph{normal} changes and changes due to camera displacement or blur/defocus. Moreover, il low-power smart cameras, tampering has to be reliably detected because false alerts would lead to unnecessary data-transmissions thus increasing the energy consumption.

@@

%What are the key components of my approach and results? Also include any specific limitations.
We here propose an algorithm to detect both camera defocus and of displacement, as stated in Section \ref{sec:probForm}. The algorithm relies on simple indicators that are computed with a low computational complexity (Section \ref{subsec:Indicators}) and leverages an outlier detection technique to detect frames yielding anomalous indicators as related to a tampering event (Section \ref{subsec:MonitoringScheme}). In particular, we perform a preliminarily partitioning of the scene into nonoverlapping regions, and separately compute and monitor indicators over different regions. Scene partitioning is defined by clustering feature vectors gathering the indicators in different pixels, thus performing a very coarse segmentation of the image (Section \ref{subsec:Segmentation}). Regions are computed during the initial configuration of the algorithm and as such do not implies any computational overhead w.r.t. operating on the whole image. Our experiments in Section \ref{sec:experiments} show that regions substantially improve the detection of camera displacements,  while when monitoring indicators meant to detect blur/defocus, it is more convenient to analyze the whole scene. Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.

@@

\subsection{Related Works}\label{subsec:relWorks}
% Why hasn't it been solved before? (Or, what's wrong with previous proposed solutions? How does mine differ?)
The literature concerning tampering detection is mostly focused on video surveillance applications and operates at few frames per second \gi{Adriano, \'e vero?}. Background models are typically leveraged to identify defocus and occlusions; in particular  ~\cite{aksay2007camera} performs detect defocus by analyzing the wavelet domain of each frame and performs histogram comparison for detecting occlusions, while camera displacements are not considered. A background-subtraction technique is employed in~\cite{saglam2009real} to identify defocus, occlusions, and displacements. \gi{FIXME Comparison in the Fourier domain for defocus detection, histogram comparison for occlusion detection, comparison between current background and delayed background for displacement detection}. Background subtraction is also used in \cite{gil2007automatic} to detect defocus, occlusions, and displacements. \gi{FIXME Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection}.
In contrast, no background models are used in \cite{alippi2010detecting}, where a sequential monitoring scheme based on a change-detection test is employed to detect changes in the average gradient energy of each frame to detect defocus due to external disturbances on the camera lens.  
\cite{ribnick2006real}: comparison between frames belonging to a buffer in order to find high values of dissimilarity, associated to tampering.
\cite{kryjak2012fpga}: implementation in a FPGA of a solution based on background modeling, histograms comparisons, edges comparisons.
\cite{harasse2004automated}: tampering detection inside a moving vehicle; uses background subtraction methods in order to identify defocus, occlusions, and displacements. Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection.
\cite{tsesmelis2013tamper}: monitoring of the number of key points extracted by SURF in order to detect defocus events, partition in blocks and HOG descriptors matching for each block in order to detect occlusions. These types of solutions requires a lot of computations

@@

Let $z_t$ be the frame acquired at time $t$, which we model as
\begin{equation}
\label{eq:observationModel}
z_t(x)=\mathcal{D}_t[y_t](x) \quad \forall x \in X
\end{equation}
where $\mathcal{D}_t$ denotes the degradation operator that transforms the original image $y_t$ in the frame $z_t$; $X \subset \mathbb{Z}^2$ denotes the regular pixels grid and $x\in \mathbb{Z}^2$ indicates the pixel coordinates. As far as there is no tampering attacks/events,
\begin{equation}
\label{eq:no_tampering}
\mathcal{D}_t[y_t](x) = y_t(x) + \eta_t(x) \quad \forall x \in X
\end{equation}
where $\eta_t$ is a random variable accounting for different noise sources (e.g., thermal, quantization, photon-counting). In normal conditions, all the images $y_t$ (thus also the frames $z_t$) might show different content but are acquired from the same viewpoint and the same camera orientation.

When at time $\tau^*$ an external disturbance introduces \emph{blur/defocus}, the image $y_t$ is degraded by a spatially variant blur operator, and $z_t$ becomes
\begin{equation}
\label{eq:model_defocus}
\mathcal{D}_t[y_t](x) = \int_{\mathcal{X}}y(s)h_t(x,s)ds\, + \eta_t(x) \quad \forall x \in X, t \geq \tau^*
\end{equation}
%where  $\mathcal{B}_t$ corresponds to a linear operator
%\begin{equation}
%\label{eq:blur}
%\mathcal{B}[y](x) = \int_{\mathcal{X}}y(s)h(x,s)ds\,,
%\end{equation}
where $h_t(x,\cdot) > 0$ is the point-spread function at pixel $x \in X$.
%\begin{equation}
%\label{eq:blur_convolution}
%\mathcal{B}[y](x) = \int_{\mathcal{X}}y(s)h(s-x)ds,
%\end{equation}
%
%\begin{equation}
%\label{eq:blur_multi}
%z_t(x)=\mathcal{D}_t[y_t](x) = \mathcal{B}_t[y_t](x) + \eta(x), \qquad x \in \mathcal{X}.
%\end{equation}
A \emph{camera displacement} at frame $\tau^*$ is instead modeled as 
\begin{equation}
\label{eq:model_displacement}
z_t(x)  = \left\{ \begin{array}{rcl}
y_t(x) + \eta(x) & \mbox{per} & t < T^* \\
w_t(x) + \eta(x) & \mbox{per} & t \geqslant T^*
\end{array}\right. ,
\end{equation}
where $w_t$ relates to a different viewpoint and/or camera orientation than $y_t$.

The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ to detect time $\tau^*$ when any tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement} occurs.
